---
title: "Baseball"
author: "Alan T. Arnholt"
output: 
  bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE, fig.align = "center")
```

Create the data set discussed on page 248 of [*OpenIntro Statistics*](https://drive.google.com/file/d/0B-DHaDEbiOGkc1RycUtIcUtIelE/view).

```{r, label = "BBsetup"}
library(dplyr)
library(openintro)
bat10 <- mlbBat10 %>%
  filter(AB > 200)
# Note unused factor levels
table(bat10$position)
bat10$position <- droplevels(bat10$position)
table(bat10$position)
# Combine 1B, 2B, 3B, SS into infield (IF)
#
bat10$pos <- plyr::revalue(bat10$position, 
                       replace = c("1B" = "IF", "2B" = "IF", "3B" = "IF", "SS" = "IF"))
table(bat10$pos)
# Note must detach plyr for dplyr to work properly
# detach("package:plyr", unload = TRUE)
```

Create the summary statistics presented in Table 5.24 of [*OpenIntro Statistics*](https://drive.google.com/file/d/0B-DHaDEbiOGkc1RycUtIcUtIelE/view).

```{r}
T524 <- bat10 %>%
  group_by(pos) %>%
  summarise(n = n(), Mean = round(mean(OBP), 4), SD = round(sd(OBP), 4))
T524
```

```{r}
DT::datatable(T524)
```

## Hypotheses {-}

**Your Turn:** 

1.  Write the null and alternative hypotheses to test whether the on base percentage is related to position.

$H_0:$

$H_A:$

2. List the conditions necessary for performing ANOVA:
  +
  +
  +




```{r, label = "BPOI", fig.cap = "Side-by-side boxplots of on base percentage according to position"}
library(ggplot2)
ggplot(data = bat10, aes(x = pos, y = OBP, fill = pos)) + 
  geom_boxplot() + 
  theme_bw() + 
  labs(y = "On base percentage", x = "Position") + 
  guides(fill = FALSE)
```

```{r, label = "QQ", fig.width = 10}
ggplot(data = bat10, aes(sample = OBP, color = pos)) + 
  stat_qq() + 
  theme_bw()
ggplot(data = bat10, aes(sample = OBP, color = pos)) + 
  stat_qq() +
  facet_grid(.~pos) +
  theme_bw() + 
  guides(color = FALSE)
```

```{r}
mod.aov <- aov(OBP ~ pos, data = bat10)
summary(mod.aov)
```

**Your turn:**  Compute the $p-$value using `pf`.  Does your answer agree with the output from the ANOVA table?

```{r}
# Your code here
```

```{r}
library(ggplot2)
ggplot(data = data.frame(x = seq(0, 6, length= 200)), aes(x = x)) +
  stat_function(fun = df, args = list(3, 233), geom = "area", fill = "purple", alpha = 0.5) +
  theme_bw() + 
  labs(x = "", y = "")

df_limit <- function(x){
  y <- df(x, 3, 323)
  y[x < 1.994] <- NA
  return(y)
}
p <- ggplot(data.frame(x = c(0, 6)), aes(x = x))
p + stat_function(fun = df_limit, geom = "area", fill = "purple", alpha = 0.4, n = 500) + 
  stat_function(fun = df, args = list(3, 323)) +
  theme_bw() +
  labs(x = "", y = "")
```

**Finished - Now consider a different problem.**

Use `bat10` to test whether the on base percentage is related to position.


```{r}
T5N <- bat10 %>%
  group_by(position) %>%
  summarise(n = n(), Mean = round(mean(OBP), 4), SD = round(sd(OBP), 4))
T5N
```

**Your turn:** 

1. Write the null and alternative hypotheses to test whether the on base percentage is related to position.

$H_O:$

$H_A:$

2. List the conditions necessary for performing ANOVA:
  +
  +
  +


```{r, , fig.width = 10}
ggplot(data = bat10, aes(x = position, y = OBP, fill = position)) + 
  geom_boxplot() + 
  theme_bw() + 
  guides(fill = FALSE)
ggplot(data = bat10, aes(sample = OBP, color = position)) + 
  stat_qq() + 
  theme_bw()
ggplot(data = bat10, aes(sample = OBP, color = position)) + 
  stat_qq() +
  facet_grid(.~position) +
  theme_bw() + 
  guides(color = FALSE)
```

```{r}
mod2.aov <- aov(OBP ~ position, data = bat10)
summary(mod2.aov)
```

**Your turn:** 

1.  Do you reject the null hypothesis?  If so, what does this mean in plain English?
2.  Which positions have a higher on base percentage?


## Bonferroni Method {-}

$$CI_{1 - \alpha_e}(\mu_i - \mu_j) = (\bar{Y}_{i\bullet} - \bar{Y}_{j\bullet}) \pm t_{1 - \tfrac{\alpha_c}{2K};df_{Error}}\sqrt{MS_{Error}}\sqrt{\frac{1}{n_i} + \frac{1}{n_j}}$$
```{r}
T5N
(0.3554 - 0.3348) +c(-1, 1)*qt(1 - 0.05/(2*choose(7,2)), 320)*sqrt(0.001182)*sqrt(1/33 + 1/39)
```

**Your Turn:**

Find a 95% Bonferroni CI for the $\mu_{3B} - \mu_{1B}$

```{r}
# Your code here
```

Another approach is to use `pairwise.t.test()`

```{r}
pairwise.t.test(bat10$OBP, bat10$position, p.adj = "bonferroni")
```


##Tukey's Honestly Significant Difference (HSD) {-}

$$CI_{1 - \alpha_e}(\mu_i - \mu_j) = (\bar{Y}_{i\bullet} - \bar{Y}_{j\bullet}) \pm \frac{q_{1 - \alpha_e:a, \nu}}{\sqrt{2}}\sqrt{MS_{Error}}\sqrt{\frac{1}{n_i} + \frac{1}{n_j}}$$
Note: $a=$ the number of treatments, $\nu=$ the degrees of freedom for $MS_{Error}$.  The notation $q_{1 - \alpha;a,\nu}$ denotes the studentized range value with $1 - \alpha$ area to the left with $a$ and $\nu$ degrees of freedom, respectively.  For example, $q_{0.95;4,20} = `r qtukey(0.95, 4, 20)`$ is obtained by entering:

```{r}
qtukey(0.95, 4, 20)
```

**Your Turn:**

Verify using the formula that $CI_{0.95}(\mu_{3B} - \mu_{1B}) = [-0.05696, -0.00899]$.

```{r}
# Your code here
```

```{r, fig.width = 8, fig.height = 8}
TukeyHSD(mod2.aov)
par(las = 1)
plot(TukeyHSD(mod2.aov))
par(las = 0)
```

-----

## Confidence Intervals (2 Groups) {-}

```{r}
NDF <- bat10 %>%
  filter(team == "BAL" | team == "BOS") %>%
  select(OBP, team) 
NDF %>%
  group_by(team) %>%
  summarize(Mean = mean(OBP), SD = sd(OBP), n = n())
ggplot(data = NDF, aes(x = team, y = OBP, fill = team)) + 
  geom_boxplot() + 
  theme_bw() + 
  labs(y = "On base percentage") + 
  guides(fill = FALSE)
ggplot(data = NDF, aes(sample = OBP, color = team)) + 
  stat_qq() + 
  theme_bw()
# Or
ggplot(data = NDF, aes(sample = OBP, color = team)) + 
  stat_qq() + 
  theme_bw() + 
  facet_grid(. ~ team)
# Standard 90% CI first
t.test(OBP ~ team, data = NDF, conf.level = 0.90)
df <- t.test(OBP ~ team, data = NDF, conf.level = 0.90)$parameter
df
```

## Bootstrap {-}

For more on the bootstrap method see [http://stat-ata-asu.github.io/STT3851ClassRepo/Rmarkdown/TheBootstrap.html](http://stat-ata-asu.github.io/STT3851ClassRepo/Rmarkdown/TheBootstrap.html).

```{r}
BALOBP <- bat10$OBP[bat10$team == "BAL"]
BOSOBP <- bat10$OBP[bat10$team == "BOS"]
obsdiff <- mean(BALOBP) - mean(BOSOBP)
obsdiff
SIMS <- 10^4 - 1
diffmean <- numeric(SIMS)
for(i in 1:SIMS){
  sampBAL <- sample(BALOBP, size = sum(!is.na(BALOBP)), replace = TRUE)
  sampBOS <- sample(BOSOBP, size = sum(!is.na(BOSOBP)), replace = TRUE)
  diffmean[i] <- mean(sampBAL) - mean(sampBOS)
}
hist(diffmean)
# OR
ggplot(data = data.frame(x = diffmean), aes(x = x)) + 
  geom_density(fill = "pink") + 
  theme_bw() + 
  labs(x = substitute(paste(bar(x)[Bal],"*", - bar(x)[Bos],"*")))
```

## Percentile Approach {-}

```{r}
BSCI <- quantile(diffmean, probs = c(0.05, 0.95))
BSCI
```

## Standard Bootstrap {-}

```{r}
SBS <- obsdiff + c(-1, 1)*qt(0.95, df)*sd(diffmean)
SBS
```



