---
title: "Inference"
author: "Alan T. Arnholt"
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE)
```

```{r}
library(ggplot2)
library(statsr)
library(openintro)
library(DT)
datatable(textbooks)
```

Is there a difference in price for textbooks between UCLA and amazon?

Graph statistic of interest first.  Note the assumption to use a paired $t-$test is that the population of differences follows a normal distribution.  This is not really an appropriate example.

```{r}
ggplot(data = textbooks, aes(x = diff)) + 
  geom_histogram(binwidth = 3, fill = "pink") + 
  theme_bw()

ggplot(data = textbooks, aes(sample = diff)) + 
  stat_qq() + 
  theme_bw()
```

Go through hypotheses

```{r}
t.test(x = textbooks$uclaNew, y = textbooks$amazNew, paired = TRUE, mu = 0)
t.test(x = textbooks$diff, mu = 0)
inference(y = diff, data = textbooks, type = "ht", statistic = "mean", method = "theoretical", null = 0, alternative = "twosided")
#
inference(y = diff, data = textbooks, type = "ht", statistic = "mean", method = "simulation", null = 0, alternative = "twosided")
```

What is wrong with the conclusion on page 230?

## Difference of two means

### Example 5.18


```{r}
datatable(ncbirths)
```

Go through hypotheses

```{r}
t.test(weight ~ habit, data = ncbirths)
#
inference(y = weight, x = habit, data = ncbirths, type = "ci", statistic = "mean", method = "theoretical", null = 0)
```

## Two versions of a course exam

```{r}
library(PASWR2)
tsum.test(mean.x = 79.4, s.x = 14, n.x = 30, mean.y = 74.1, s.y = 20, n.y = 27, alternative = "greater")
```
